---
title: "QUIZ"
author: "Fabian Abrego"
date: "4/27/2020"
output:
  pdf_document: default
  html_document: default
---

#Step Wise Model Selection 

##Question

For this question, we will be using the `teengam` data set included in the `Faraway` library. 


**(a)** Set `gamble` as your response variable and using stepwise model selection, derive the best model. Do not use the `step` function for this. Use the `extarctAIC` function for AIC values instead of the `AIC` function. 


```{r}
gamble = faraway::teengamb
modelstart = lm(gamble ~ 1, data = gamble)


#First step 
extractAIC(modelstart) #325.3354 Starting AIC 

#Add a variable
extractAIC(lm(gamble ~ sex, data = gamble)) #318.7866 -  Improvement 
extractAIC(lm(gamble ~ status, data = gamble)) #327.2157 - No Improvement 
extractAIC(lm(gamble ~ income, data = gamble)) #304.3356 - Greatest Improvement - Add to model
extractAIC(lm(gamble ~ verbal, data = gamble)) #325.0025 - Improvement 

#Second Step 
#Current AIC = 304.3356

#Add a variable
extractAIC(lm(gamble ~ income + sex, data = gamble)) #296.6268 - Add to model 
extractAIC(lm(gamble ~ income + status, data = gamble)) #305.112 - No Improvement 
extractAIC(lm(gamble ~ income + verbal, data = gamble)) #305.3536 - No Improvement 

#Remove a variable
extractAIC(modelstart) #325.3354 - No Improvement 

#Third step 
#Current AIC = 296.6268

#Add a variable 
extractAIC(lm(gamble ~ income + sex + status, data = gamble)) #298.2086 - No Improvement 
extractAIC(lm(gamble ~ income + sex + verbal, data = gamble)) #296.2145 - Add to model 

#Remove a variable 
extractAIC(lm(gamble ~ income, data = gamble)) #304.3356 - No Improvement 
extractAIC(lm(gamble ~ sex, data = gamble)) #318.7866 - No Improvement 

#Fourth Step 
#Current AIC = 296.2145

#Add a variable 
extractAIC(lm(gamble ~ income + sex + verbal + status, data = gamble)) #298.1758 - No Improvement 

#Remove a variable 
extractAIC(lm(gamble ~ income + sex, data = gamble)) #296.6268 - No Improvement 
extractAIC(lm(gamble ~ income + verbal, data = gamble)) #305.3536 - No Improvement 
extractAIC(lm(gamble ~ sex + verbal, data = gamble)) #316.6443 - No Improvement 

Final_Model = lm(gamble ~ income + sex + verbal, data = gamble)
```

**(b)** Use the `step` function to derive the best model from stepwise selection. Are they the same?

```{r}
#Using Step Function
model_step = step(modelstart, scope = gamble ~ sex + status + income + verbal, direction = "both")

#They are equal
summary(model_step)
summary(Final_Model)
```


The inspiration for this problem was on the basis of wanting to understand the stepwise selection process as when it was first presented, it left me confused in regard to the logic of the process. The problem and solution was all worked out and did not contribute outside sources into its creation. 
